---
title: "Slide 01: Data"
author: "Richard Dien Giau Bui"
format: revealjs
editor: visual
---

## Load packages

-   First, before a code session, think which packages you need to use today

    -   Like before cooking, which knifes we should use

-   Load it with `library(name_of_package)`:

```{r}
#| echo: true
#| warning: false
library(tidyverse)
library(skimr)
```

## Code block

-   Insert a new code block with **CTRL+ALT+I** then run with the green **RUN** button

```{r}
#| echo: true
60 + 8
```

## Four main data types in R

-   `numeric`: for example, counting, km, scores, money, ...

-   `character`: for example, "Wednesday", "Today I failed the coding class", ...

-   `factor`: e.g., a column with only 4 possible categories: A, B, C, D

-   `date`: `2024-02-28`

## Type 1. `numeric` 

-   Create a numeric variable

```{r}
#| echo: true
x = c(1:10) # other way: x <- c(1:10)
typeof(x)
x = c(0.1, 0.4, 0.5)
typeof(x)
```

-   Some math manipulation with numeric

```{r}
#| echo: true
length(x)
mean(x)
median(x)
min(x)
max(x)
```

-   Quick way:

```{r}
#| echo: true
#| eval: false
skim(x)
```

## Type 2: `character`

-   For text data

-   Can't use with math calculation, for example, below code will raise error:

```{r}
#| echo: true
#| eval: false
"60" + "8"
```

-   Create text vector

```{r}
#| echo: true
s = c("ann", "betty", "candy")
```

-   Some text manipulation with `stringr` package:

```{r}
#| echo: true
length(s)
str_length(s)
str_to_upper(s)
str_replace(string = s, pattern = "a", replacement = "$")
paste0("student_", s)
```

## Type 3: `factor`

-   Some category only, not continuous numeric variable

-   For example:

```{r}
#| echo: true
f1 = c(0, 1, 1, 0 ,1)
f2 = c("Monday", "Tuesday", "Wednesday", "Tuesday", "Friday")
```

-   Sometimes, we need `factor` to do some special manipulation, e.g., plotting data by groups

## Type 4: `date`

-   Make date:

```{r}
#| echo: true
d = c(20211231, 20221231, 20231231)
typeof(d)
```

- Convert `numeric` in yyyymmdd format to `date`:
```{r}
#| echo: true
d = ymd(d)
typeof(d)
```


